---
title: "PrEP Optimization Summary"
author: "Greg Knowlton, Emeli Anderson, Sam Jenness"
date: "7/13/2020"
output: html_document
---

The central question is: how should we allocate a fixed budget to three PrEP support interventions to maximize the number of HIV infections averted over the course of 10 years among MSM in Atlanta?

The policy levers being considered in this model:
•	POIP – percentage of MSM reached by initiation intervention
•	POAC_yr – adherence intervention capacity (max. number of people who can receive the intervention per year)
•	PORC – retention intervention capacity (max. number of people who can be active in the intervention in any given week)

1.	Estimate cost per unit of capacity for each of the initiation, adherence, and retention interventions. 
2.	Fit meta-model to map program capacities (POIP, POAC_yr, PORC) to infections averted. 
    e.g.  InfAvert = f(POIP, POAC_yr, PORC).
3.	Use an optimization algorithm to identify the optimal allocation of a fixed budget, B.
4.	Explore how the optimal allocation changes with different budget constraints.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Required packages
library(dplyr)
library(psych)
library(ggplot2)
library(tidyverse)
# package for generalized additive models (gam)
library(mgcv)

# Import optimization results
load("C:/Users/Greg/Desktop/PrEP-Optimize/optim_res.rda")
# Import simulation output
df_prep <- readRDS("C:/Users/Greg/Desktop/PrEP-Optimize/analysis/data/prepOptim-Yearly-v4-1000sets-250per.rds")

# Linear costs for PrEP support program capacities
cost.i <- 10 * (80 * 790 * 100)
cost.a <- 1 * (511.38 * 10)
cost.r <- 0.33 * 10 * (77.17/(52 * 337/365) + (50.17 * (52/12)))
```

```{r data_cleaning, include = FALSE}
# translate weekly adherence capacity into yearly capacity
df_prep$POAC_yr <- df_prep$POAC*52

# calculate 10-year outcomes
df_prep_10yr <- full_join(df_prep %>% group_by(scenario) %>%
                            summarise_at(vars(incid, infAvert, PrEPStarts),
                                         sum),
                          df_prep %>% group_by(scenario) %>%
                            summarise_at(vars(PrEPHighAdr, prepElig, prepCurr, pCov, PrEPHighAdr),
                                         mean),
                          by="scenario")

df_prep_10yr <- full_join(df_prep_10yr,
                          df_prep %>% group_by(scenario) %>%
                            summarise_at(vars(POIP, PORC, POAC_yr),
                                         first),
                          by = "scenario")

df_prep_10yr <- full_join(df_prep_10yr,
                          df_prep %>% group_by(scenario) %>%
                            summarise_at(vars(OptimInitStarts, OptimAdhrStarts, OptimRetnStarts),
                                         sum),
                          by="scenario")
```

Simulation Output Visualization:

The outcome we are looking at here is the proportion of active PrEP users who are highly adherent. We expect this proportion to go up with increasing adherence intervention capacity. At a certain point, this highly adherent proportion stops increasing with adherence capacity because all new PrEP users are already enrolled in the adherence intervention. When the uptake intervention is well-funded (High PrEP Uptake), a greater adherence capacity can be utilized. Consequently, we see the proportion of active PrEP users who are highly adherent start to plateau at a lower adherence capacity when PrEP uptake is also low.

```{r high_adherence_plot}

df_prep_10yr$POIP_cat <- cut(df_prep_10yr$POIP, quantile(df_prep_10yr$POIP, c(0, .25, .75, 1)), labels = c("Low PrEP Uptake", "Medium PrEP Uptake", "High PrEP Uptake"), include.lowest = TRUE)

ggplot(df_prep_10yr, aes(x = POAC_yr, y = PrEPHighAdr, color = POIP_cat)) + geom_jitter() + scale_color_manual(values = c("#D55E00", "#999999", "#009E73"))
```
The y-axis denotes the mean PrEP coverage proportion for each simulation over the course of the 10 year time horizon. Increasing the rate of PrEP uptake and the average length of time each PrEP user stays on the medication will both raise overall coverage levels. Increasing retention intervention capacity raises coverage levels until the capacity exceeds the number of PrEP users who can utilize the capacity. Increasing PrEP uptake shifts this point of excessive retention capacity to a higher value. Visually speaking, it appears that marginal effect of retention capacity on PrEP coverage is the same at all levels of PrEP uptake, except in the space where retention capacity exceeds demand. If infections averted is a function of PrEP coverage, then this suggests that the lack of a synergistic interaction between the uptake and retention interventions when retention demand is not exceeded.

```{r pCov_porc_plot}
ggplot(df_prep_10yr, aes(x = PORC, y = pCov, color = POIP_cat)) + geom_jitter() + scale_color_manual(values = c("#D55E00", "#999999", "#009E73"))
```
This figure shows the same dynamics in a slightly different way. The higher retention capacity values don't lead to an increase in PrEP coverage, and the relationship between coverage levels and POIP appears to be roughly linear and independent of retention capacity. 

```{r pCov_poip_plot}
# POIP has a similar range of effects on pCov as PORC
# Most PORC values are much higher than prevalent PrEP users, leading to thick band
df_prep_10yr$PORC_cat <- cut(df_prep_10yr$PORC, quantile(df_prep_10yr$PORC, c(0, .25, .75, 1)), labels = c("Low Retention Capacity", "Medium Retention", "High Retention Capacity"), include.lowest = TRUE)
ggplot(df_prep_10yr, aes(x = POIP, y = pCov, color = PORC_cat)) + geom_jitter() + scale_color_manual(values = c("#D55E00", "#999999", "#009E73"))
```
We use a generalized additive model (GAM) to fit a flexible and smooth function to our simulation output. In this statistical modeling approach, our outcome is cumulative infections averted, and the predictors of interest are uptake program initiation proportion (POIP), retention program capacity (PORC), and adherence program capacity (POAC_yr). Like generalized linear models, GAMs are specified by a link function and a mean variance relationship (family). Because our outcome (infections averted) is positive and continuous, we used a Gamma family mean-variance relationship, and we chose to use a log link function because it provided the best fit to our data. In GAMs, relationships between the individual predictors and the dependent variable follow smooth patterns that can be linear or nonlinear. 

```{r gam_fit}
gam <- gam(data = df_prep_10yr,
           formula = infAvert ~ s(POIP, k = 4) + s(POAC_yr, k = 4) + s(PORC, k = 4) + ti(POIP, PORC, k = 4) + ti(POIP, POAC_yr, k = 4) + ti(PORC, POAC_yr, k = 4),
           family = Gamma(link = "log"))
```


We can visualize the generalized additive model predictions for infections averted to get a sense of how the program parameters interact and affect outcomes. On these plots, "response" is infections averted, and on the three dimensional plots, the axes span the entire allowable range of each capacity parameter (POIP: 0-0.10, POAC_yr: 0-2000, PORC: 0-2000). 

Uptake/Retention Interaction plots:

```{r vis_gam}
vis.gam(gam, view = c("PORC", "POIP"), type = "response", theta = -45)
vis.gam(gam, view = c("PORC", "POIP"), plot.type = "contour", type = "response")
```

Uptake/Adherence Interaction plots:

```{r}
vis.gam(gam, view = c("POAC_yr", "POIP"), type = "response", theta = -45)
vis.gam(gam, view = c("POAC_yr", "POIP"), plot.type = "contour", type = "response")
```

Adherence/Retention Interaction plots:

```{r}
vis.gam(gam, view = c("POAC_yr", "PORC"), type = "response", theta = -45)
vis.gam(gam, view = c("POAC_yr", "PORC"), plot.type = "contour", type = "response")
```

The y-axis denotes the predicted number of infections averted based on the generalized additive model and the three different covariate values. The plot has three facets for the minimum, median, and maximum values for the uptake intervention parameter, and each curve is color-coded according the minimum, median, and maximum adherence intervention capacity. This allows one to visualize the effect of changing retention capacity on infections averted at various combinations of the other program capacities.

```{r}
n = 101
grid.df <- expand.grid(POIP = seq(0, 0.10, length.out = n),
                       POAC_yr = seq(0, 2000, length.out = n),
                       PORC = seq(0, 2000, length.out = n))
pred <- exp(predict(gam,
                newdata = grid.df))
pred.df <- cbind(pred, grid.df)

porc <- pred.df %>% filter((POIP == min(POIP) | POIP == max(POIP) | POIP == median(POIP)) &
                             (POAC_yr == min(POAC_yr) | POAC_yr == max(POAC_yr) | POAC_yr == median(POAC_yr)))

ggplot(porc, aes(x = PORC, y = pred, color = POAC_yr)) + geom_point() + facet_wrap(.~as.factor(POIP))
```

These plots are faceted by uptake intervention capacity and color coded according to retention intervention capacity.


```{r}
poac <- pred.df %>% filter((POIP == min(POIP) | POIP == max(POIP) | POIP == median(POIP)) &
                             (PORC == min(PORC) | PORC == max(PORC) | PORC == median(PORC)))

ggplot(poac, aes(x = POAC_yr, y = pred, color = PORC)) + geom_point() + facet_wrap(.~as.factor(POIP))
```

These plots are faceted by adherence intervention capacity, and color coded by retention intervention capacity.


```{r}
poip <- pred.df %>% filter((PORC == min(PORC) | PORC == max(PORC) | PORC == median(PORC)) &
                             (POAC_yr == min(POAC_yr) | POAC_yr == max(POAC_yr) | POAC_yr == median(POAC_yr)))

ggplot(poip, aes(x = POIP, y = pred, color = PORC)) + geom_point() + facet_wrap(.~as.factor(POAC_yr))
```

The plots above demonstrate how shifts in the intervention capacity parameters influence infections averted, but we have not yet considered the unit costs for increasing intervention capacity. The following unit costs are assumed for the budget optimization problem:

~$632,000 per percentage point increase to PrEP Optimization Initiation Proportion
~$5,100 per yearly adherence intervention capacity slot
~$720 per retention intervention capacity slot (each slot used for an average of 337 days)

Budget optimization results for variable budget sizes:

```{r}
res_plot <- res %>%
  filter(converge == 0) %>%
  rowwise() %>%
  mutate(poip_budget_prop = poip * cost.i / budget,
         poac_budget_prop = poac * cost.a / budget,
         porc_budget_prop = porc * cost.r / budget,
         poip_budget = poip * cost.i,
         poac_budget = poac * cost.a,
         porc_budget = porc * cost.r)
# plots showing how optimal poip, poac, pocr, and infAvert change with budget constraint
ggplot(data = res_plot, aes(x = budget, y = infAvert)) + geom_area()
```

```{r}
ggplot(data = res_plot, aes(x = budget, y = poip)) + geom_area()
ggplot(data = res_plot, aes(x = budget, y = poac)) + geom_area()
ggplot(data = res_plot, aes(x = budget, y = porc)) + geom_area()
```


```{r}
# plots showing what fraction of budget is allocated to each intervention as a function of budget constraint
res_plot_area_prop <- res_plot %>%
  pivot_longer(cols = c(poip_budget_prop, poac_budget_prop, porc_budget_prop),
               names_to = "program")

ggplot(res_plot_area_prop, aes(x = budget, y = value, fill = program)) +
  geom_area()
```

```{r}
res_plot_area <- res_plot %>%
  pivot_longer(cols = c(poip_budget, poac_budget, porc_budget),
               names_to = "program")
ggplot(res_plot_area, aes(x = budget, y = value, fill = program)) +
  geom_area()
```



